---
title: "Milestone_1"
author: "Hanna Zdulska"
date: "3/14/2020"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Zbiór

## Opis
Context:
The data were obtained in a survey of students math and portuguese language courses in secondary school. It contains a lot of interesting social, gender and study information about students. You can use it for some EDA or try to predict students final grade.

Content:
Attributes for both student-mat.csv (Math course) and student-por.csv (Portuguese language course) datasets:

school - student's school (binary: 'GP' - Gabriel Pereira or 'MS' - Mousinho da Silveira)
sex - student's sex (binary: 'F' - female or 'M' - male)
age - student's age (numeric: from 15 to 22)
address - student's home address type (binary: 'U' - urban or 'R' - rural)
famsize - family size (binary: 'LE3' - less or equal to 3 or 'GT3' - greater than 3)
Pstatus - parent's cohabitation status (binary: 'T' - living together or 'A' - apart)
Medu - mother's education (numeric: 0 - none, 1 - primary education (4th grade), 2 – 5th to 9th grade, 3 – secondary education or 4 – higher education)
Fedu - father's education (numeric: 0 - none, 1 - primary education (4th grade), 2 – 5th to 9th grade, 3 – secondary education or 4 – higher education)
Mjob - mother's job (nominal: 'teacher', 'health' care related, civil 'services' (e.g. administrative or police), 'at_home' or 'other')
Fjob - father's job (nominal: 'teacher', 'health' care related, civil 'services' (e.g. administrative or police), 'at_home' or 'other')
reason - reason to choose this school (nominal: close to 'home', school 'reputation', 'course' preference or 'other')
guardian - student's guardian (nominal: 'mother', 'father' or 'other')
traveltime - home to school travel time (numeric: 1 - 1 hour)
studytime - weekly study time (numeric: 1 - 10 hours)
failures - number of past class failures (numeric: n if 1<=n<3, else 4)
schoolsup - extra educational support (binary: yes or no)
famsup - family educational support (binary: yes or no)
paid - extra paid classes within the course subject (Math or Portuguese) (binary: yes or no)
activities - extra-curricular activities (binary: yes or no)
nursery - attended nursery school (binary: yes or no)
higher - wants to take higher education (binary: yes or no)
internet - Internet access at home (binary: yes or no)
romantic - with a romantic relationship (binary: yes or no)
famrel - quality of family relationships (numeric: from 1 - very bad to 5 - excellent)
freetime - free time after school (numeric: from 1 - very low to 5 - very high)
goout - going out with friends (numeric: from 1 - very low to 5 - very high)
Dalc - workday alcohol consumption (numeric: from 1 - very low to 5 - very high)
Walc - weekend alcohol consumption (numeric: from 1 - very low to 5 - very high)
health - current health status (numeric: from 1 - very bad to 5 - very good)
absences - number of school absences (numeric: from 0 to 93)
These grades are related with the course subject, Math or Portuguese:

G1 - first period grade (numeric: from 0 to 20)
G2 - second period grade (numeric: from 0 to 20)
G3 - final grade (numeric: from 0 to 20, output target)
Additional note: there are several (382) students that belong to both datasets .
These students can be identified by searching for identical attributes
that characterize each student, as shown in the annexed R file.

Source Information
P. Cortez and A. Silva. Using Data Mining to Predict Secondary School Student Performance. In A. Brito and J. Teixeira Eds., Proceedings of 5th FUture BUsiness TEChnology Conference (FUBUTEC 2008) pp. 5-12, Porto, Portugal, April, 2008, EUROSIS, ISBN 978-9077381-39-7.

Fabio Pagnotta, Hossain Mohammad Amran.
Email:fabio.pagnotta\@studenti.unicam.it, mohammadamra.hossain '@' studenti.unicam.it
University Of Camerino

```{r }
options(stringsAsFactors = F)
library(DataExplorer)
library(ggplot2)
library(patchwork)
library(dplyr)
df <- read.csv("./students-all.csv")
df <- df %>% select(-X)

# to make procentowa tabelka
ptable <- function(s1, s2){
  tab <- table(df[,c(s1,s2)])
  as.data.frame(prop.table(tab, 1)*100)
}
```

## 

You can also embed plots, for example:

```{r}
slices <- which(colnames(df)  %in% c("age", "failures", "absences", "G1", "G2", "G3"))
df[, -slices] <- lapply(df[, -slices], as.factor)
head(df)
summary(df)
```

```{r}
DataExplorer::plot_bar(df)
```


```{r}
DataExplorer::plot_histogram(df)
```



```{r}
DataExplorer::plot_density(df)
```
```{r}
DataExplorer::plot_boxplot(df, by = "romantic")
```


```{r}
DataExplorer::plot_boxplot(df, by = "romantic")
```

```{r}
# to nie wiem czy zatrzymywać
# kto ile studjuje/podróżuje
ggplot(df, aes(x = traveltime, y=studytime, color = romantic)) + geom_jitter()
# raczej nie widać nic specjalnie

ggplot(df, aes(x = Dalc, y=Walc, color = romantic)) + geom_jitter()
# tendencja do picia w weekendy, ale wytrywali romantycy w [4,5] Dalc

# rozkład wyników w zależności od związku
ggplot(df, aes(x = G1, color = romantic)) + geom_density()
ggplot(df, aes(x = G2, color = romantic)) + geom_density()
ggplot(df, aes(x = G3, color = romantic)) + geom_density()
# rośnie ogon niezdawalności -> u romantyków jakby bardziej no no 


# kto pije więcej, ale to ludzie, a nie procentowo
ggplot(df, aes(x = Dalc, fill = romantic)) + geom_bar(position='dodge')
ggplot(df, aes(x = Walc, fill = romantic)) + geom_bar(position='dodge')


```
```{r}
# kto pije więcej procentowo ?
tab <- table(df[,c("romantic", "Walc")])
x <- as.data.frame(prop.table(tab, 1)*100)
ggplot(x, aes(x = Walc, y = Freq, fill = romantic)) + geom_bar(position = "dodge", stat = "identity")
# no widać, że walą w weekendy po równo

# a w normalne dni?
tab <- table(df[,c("romantic", "Dalc")])
x <- as.data.frame(prop.table(tab, 1)*100)
ggplot(x, aes(x = Dalc, y = Freq, fill = romantic)) + geom_bar(position = "dodge", stat = "identity")
# delikatnie romantycy więcej
```
```{r}
# A łącznie? wg ludzi (sumarycznie Walc + Dalc)
ggplot(df, aes(x = as.integer(Walc) + as.integer(Dalc), fill = romantic )) + geom_bar(position = "dodge")

df2 <- df %>% 
  group_by(romantic) %>%
  transmute(alc = as.integer(Dalc) + as.integer(Walc))
x <- as.data.frame(prop.table(table(df2), 1)*100)

# procentowo
ggplot(x, aes(x = alc, y = Freq, fill = romantic)) + geom_bar(position = "dodge", stat = "identity")
# no jakby romantycy więcej walili
```

```{r}
# A w zależności od kierunku? wg ludzi (sumarycznie Walc + Dalc)
ggplot(df, aes(x = as.integer(Walc) + as.integer(Dalc), fill = major )) + geom_bar(position = "dodge")

df2 <- df %>% 
  group_by(major) %>%
  transmute(alc = as.integer(Dalc) + as.integer(Walc))
x <- as.data.frame(prop.table(table(df2), 1)*100)

# procentowo
ggplot(x, aes(x = alc, y = Freq, fill = major)) + geom_bar(position = "dodge", stat = "identity")
# no równo no!
```

```{r}
# to może mają więcej nieobecności
ggplot(df, aes(x = absences, color = romantic)) + geom_density() + facet_grid(major~Walc)
# no i jakby ci bardziej pijani mieli więcej nieobecności
# ale czy to jesy wystarczająca próbka? Może jest tylko po 3 osoby pijące na lvl5?
table(df[,c("major", "Walc")])

# A w Dalc?
ggplot(df, aes(x = absences, color = romantic)) + geom_density() + facet_grid(major~Dalc)
# jakby bardziej się obijało na nieobecnościach

table(df[,c("major", "Dalc")])
# Ale to jednak mała próbka
```


```{r}
# to może gdzie mieszkają ma wpływ?
ggplot(df, aes(G3, color = romantic)) + geom_density() + facet_wrap(~address)
# no też nie do końca...

# A jak picie w dniach tygodnia ma odbicie na ocenach?
ggplot(df, aes(G1, color = romantic)) + geom_density() + facet_wrap(~Dalc)
ggplot(df, aes(G2, color = romantic)) + geom_density() + facet_wrap(~Dalc)
ggplot(df, aes(G3, color = romantic)) + geom_density() + facet_wrap(~Dalc)
# sprawdzmy próbkę
table(df[, c("Dalc", "romantic")])
```

